---
title: "Background Environmental Data"
author: "Haley Poppinga"
date: "2026/01/19"
format:  
  html:
    toc: true
    theme: sandstone
---

  
## Respo Code for Light and Dark Runs
### Created by: Haley Poppinga, Maya Powell, and Nyssa Silbiger
#### Last updated on: 2026-02-09

## Introduction
This script will help us process the raw data gathered during respirometry runs. 
Need to change for specific project/experimental variables 

### Install Package
if these packages are not yet installed, install them 
great for updates or new users 

if ("segmented" %in% rownames(installed.packages()) == 'FALSE') install.packages('segmented')  
if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix')  
if ("LoLinR" %in% rownames(installed.packages()) == 'FALSE') devtools::install_github('colin-olito/LoLinR')
if ("chron" %in% rownames(installed.packages()) == 'FALSE') install.packages('chron')
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse')
if ("here" %in% rownames(installed.packages()) == 'FALSE') install.packages('here')
if ("patchwork" %in% rownames(installed.packages()) == 'FALSE') install.packages('patchwork')
if ("PNWColors" %in% rownames(installed.packages()) == 'FALSE') install.packages('PNWColors')
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') install.packages('rTPC')

## Load Libraries
Include Versions of libraries

```{r}
#| warning: false
#| message: false

library(segmented)
library(plotrix)
library(gridExtra)
library(LoLinR)
library(lubridate)
library(chron)
library(patchwork)
library(tidyverse)
library(here)
library(PNWColors)
library(ggrepel)
library(reshape2)
library(viridis)
library(car)
library(future)
library(furrr)
library(dplyr)
```



now it's time to code
get the file path

set the path to all of the raw oxygen datasheets
these are saved onto the computer in whatever file path/naming scheme you saved things to 
```{r}
path.p<-here("Data","RespoFiles","TPC","RawO2") #the location of all your respirometry files
```

you can change to individual run folders if needed

# bring in all of the individual files
filenames_final<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE)) #list all csv file names in the folder and subfolders

#basename above removes the subdirectory name from the file, re-name as file.names.full
file.names.full<-list.files(path = path.p, pattern = "csv$", recursive = TRUE) 

#empty chamber volume
ch.vol <- 475 #mL #of small chambers 

######### Load and tidy files ###############
############################################
#Load your respiration data file, with all the times, water volumes(mL), #not doing dry weight just SA
#RespoMeta <- read_csv(here("Data","RespoFiles","Respo_Metadata_SGDDilutions_Cabral_Varari.csv"))
BioData <- read_csv(here("Data","RespoFiles","TPC","Fragment_Measurements_TPC.csv"))

RespoMeta <- read_csv(here("Data","RespoFiles","TPC","TPC_meta.csv"))
#View(BioData)
#View(RespoMeta)

# species names
sp_names <- read_csv(here("Data", "species_names.csv"))

# join the data together
BioData <- BioData %>% 
  full_join(sp_names) %>%
  dplyr::select(-full_species, -species_ID)

RespoMeta <- RespoMeta %>% 
  dplyr::select(-notes)

Sample_Info <- left_join(RespoMeta, BioData)
#View(Sample_Info)

##### Make sure times are consistent ####
# make start and stop times real times, so that we can join the respo output and sample_info data frames
Sample_Info <- Sample_Info %>% 
  #drop_na(sample_ID) %>% 
  unite(date,start_time,col="start_time",remove=F, sep=" ") %>% 
  unite(date,stop_time,col="stop_time",remove=F, sep=" ") %>%
  mutate(start_time = mdy_hms(start_time)) %>% 
  mutate(stop_time = mdy_hms(stop_time)) %>% 
  mutate(date = mdy(date))

#View(Sample_Info)
write_csv(Sample_Info, here("Data","RespoFiles","TPC","Sample_Info.csv"))
Sample_Info <- read_csv(here("Data","RespoFiles","TPC","Sample_Info.csv"))

#generate a 4 column dataframe with specific column names
# data is in umol.L.sec

n_temp_levels<-9 # number of unique temperature levels

RespoR <- tibble(.rows =length(filenames_final)*n_temp_levels*2, # temp * 2 for light and dark runs
                 sample_ID = NA,
                 Intercept = NA,
                 umol.L.sec = NA,
                 Temp.C = NA,
                 temp_c_level = NA,
                 temp_c_value = NA,
                 run_block = NA,
                 light_dark = NA)

######### Create a for loop! ###############
############################################

###forloop##### 
for(i in 1:length(filenames_final)) {
  FRow <- as.numeric(which(Sample_Info$FileID_csv==filenames_final[i])) # stringsplit this renames our file
  
  Respo.Data1 <- read_csv(skip=1,file.path(path.p, paste0(file.names.full[i]))) %>% # reads in each file in list
    dplyr::select(Date, Time, Value, Temp) %>% # keep only what we need: Time stamp per 1sec, Raw O2 value per 1sec, in situ temp per 1sec
    unite(Date,Time,col="Time",remove=T, sep = " ") %>%
    drop_na() %>% 
    mutate(Time = mdy_hms(Time)) #%>% # convert time
    #mutate(help = i) ##if stuck in forloop with error from filter, can check RespoR and see at what row the forloop stopped working  
  
  ## cut the data by start and stop times from metadata
  #Use start time of each light step from the metadata to separate data by light stop
  
  oxy_subsets <- Sample_Info[FRow,] %>%
    pmap(function(temp_c_value, light_dark, start_time, stop_time, ...) {
      data <- Respo.Data1  %>%
        filter(Time >= start_time & Time <= stop_time) %>%
        arrange(Time) %>%
        mutate(t_sec = as.numeric(difftime(Time, first(Time), units = "secs"))) %>% #keep everything in seconds
        mutate(light_dark = light_dark,
               temp_c_value = temp_c_value) %>%
        filter(t_sec > 120) %>%                          # drop first 2 min (120 s)
        filter(row_number() %% 10 == 0)                  # keep every 10th row
      # now t_sec increments by ~20 s for your kept rows
      #   filter(Time >= start_time & Time <= stop_time) %>%
      #   mutate(sec = row_number()) %>%# add an id for each row to help remove the first few mins
      #   mutate(light_dark = light_dark,
      #          temp_c_value = temp_c_value,
      #          sec = sec) %>%
      #   filter(sec > 60)  %>%# delete the first 2 mins of data assuming freq of 2 Hz
      #   mutate(row_number = row_number()) %>%
      #   filter(row_number %% 10 == 0) %>%  # keep every 10th row only to thin the data
      #   dplyr::select(-row_number) %>%
      #   mutate(sec2 = row_number())  #update the row numbers
      # #return(subset)
    }) 
  
  
  # Combine into one long dataframe with ID labels
  combined_oxy <- bind_rows(oxy_subsets)
  
  # Get the filename without the .csv
  rename<- sub("_O2.csv","", filenames_final[i])
  
  ### plot and export the thinned data ####
  p1<- ggplot(combined_oxy, aes(x = sec, y = Value)) +
    geom_point(aes(color = light_dark)) +
    labs(
      x = 'Time (seconds)',
      y = expression(paste(' O'[2],' (',mu,'mol/L)')),
      title = "original"
    )+
    facet_wrap(~temp_c_value, scales = "free_y")
  
  
  ##Olito et al. 2017: It is running a bootstrapping technique and calculating the rate based on density
  #option to add multiple outputs method= c("z", "e "pc")
  
  # Define function for fitting LoLinR regressions to be applied to all intervals for all samples
  fit_reg <- function(data) {
    rankLocReg(xall = data$t_sec, yall = data$Value, #changed to secs in time instead of sec2 variable
               alpha = 0.2, method = "pc", verbose = FALSE)
  }
  
  # Setup for parallel processing
  future::plan(multisession)
  
  # Map LoLinR function onto all intervals of each sample's thinned dataset
  df <- combined_oxy %>%
    dplyr::select(t_sec, Value, temp_c_value, Temp, light_dark)%>%
    #mutate(sec2 = as.numeric(sec2))%>%
    nest_by(temp_c_value, light_dark) %>%
    ungroup()%>%
    mutate(regs = furrr::future_map(data, fit_reg), # run the LOLinR fit in parallel
           Temp.C = map_dbl(map(data, "Temp"), mean),# get the mean temperature
           RegStats =map(regs, function(x){ # extract the intercept and slope for the parameters
             x$allRegs %>%
               slice(1) %>%
               dplyr::select(Intercept = b0,
                      umol.L.sec = b1)
           }) )
  
  
  #  Plot regression diagnostics
  df <- df %>% 
    mutate(temp_light = paste(temp_c_value,light_dark))
  
  for(j in 1:length(df$temp_light)){
    pdf(paste0(here("Output","TPC"),"/",rename,"_",j,".pdf" ))
    plot(df$regs[[j]])
    dev.off() 
  }
  
  
  df<-df %>%
    dplyr::select(temp_c_value,Temp.C,light_dark,RegStats ) %>%
    unnest(RegStats) %>%
    mutate(sample_ID = rename) %>%
    left_join(Sample_Info[FRow,] %>%
                dplyr::select(sample_ID, temp_c_level, temp_c_value, run_block, light_dark))  # make sure the light value (or whatever other metadata you want) is in the final DF
  
  ################################
  # fill in all the O2 consumption and rate data
  
  RespoR[FRow,"Temp.C"]<-df$Temp.C
  RespoR[FRow, "light_dark"]<-df$light_dark
  RespoR[FRow,"sample_ID"]<-df$sample_ID
  RespoR[FRow,"Intercept"]<-df$Intercept
  RespoR[FRow,"umol.L.sec"]<-df$umol.L.sec
  RespoR[FRow,"temp_c_level"]<-df$temp_c_level
  RespoR[FRow,"temp_c_value"]<-df$temp_c_value
  RespoR[FRow,"run_block"]<-df$run_block
  
}  

#View(RespoR)
######### end of for loop - celebrate victory of getting through that ###############
############################################

#export raw data and read back in as a failsafe 
#this allows me to not have to run the for loop again !!!!!
write_csv(RespoR, here("Data","RespoFiles","TPC","Respo_R.csv"))  
#MP ran for TPC 1-6 (all) on Sept 29th 2025
##### 

######### Calculate Respiration rate ###############
############################################

RespoR <- read_csv(here("Data","RespoFiles","TPC","Respo_R.csv"))
Sample_Info <- read_csv(here("Data","RespoFiles","TPC","Sample_Info.csv"))
ch.vol <- 475 #mL #of small chambers 

RespoR2 <- RespoR %>%
  #drop_na(FileID_csv) %>% # drop NAs
  left_join(Sample_Info) %>% # Join the raw respo calculations with the metadata
  mutate(Ch.Volume.mL = ch.vol) %>% # measured volume of chambers with coral + stand + stirbar displacement
  mutate(Ch.Volume.L = Ch.Volume.mL * 0.001) %>% # mL to L conversion
  mutate(umol.sec = umol.L.sec*Ch.Volume.L) %>% #Account for chamber volume to convert from umol L-1 s-1 to umol s-1. This standardizes across water volumes (different because of coral size) and removes per Liter
  mutate_if(sapply(., is.character), as.factor)  #convert character columns to factors

#Account for blank rate by sample run Block (if we do at least one blank per block)

#View(RespoR)

####### normalize the respo rates to the blanks ##### 

RespoR_Normalized <- RespoR2 %>% 
  filter(blank == 1) %>% # grab the blanks
  group_by(temp_c_value, run_block,light_dark) %>%
  #dplyr::select(blank.rate = umol.sec) %>% ## rename the blank column 
  summarise(blank.rate = mean(umol.sec, na.rm = TRUE)) %>% # if you have multiple blanks per run take the average
  ungroup() %>% 
  #dplyr::select(Light_level, run_block, blank.rate, date) %>% # this is what I will use to join the blanks back with the raw data
  right_join(RespoR2) %>% # join blanks with the respo data
  mutate(umol.sec.corr = umol.sec - blank.rate, # subtract the blank rates from the raw rates   
         umol.cm2.hr = (umol.sec.corr*3600)/SA_cm2,
         umol.cm2.hr_uncorr = (umol.sec*3600)/SA_cm2) %>%
         #mmol.cm2.hr = 0.001*(umol.sec.corr*3600)/SA_cm2, # convert to mmol cm-2 hr-1
         #mmol.cm2.hr_uncorr = 0.001*(umol.sec*3600)/SA_cm2) %>% 
  filter(blank == 0) %>% # remove the Blank data
  #ungroup() %>%
  dplyr::select(date, species, sample_ID, frag_ID, light_dark, run_block, SA_cm2, run_block, umol.cm2.hr, chamber_channel, 
                Temp.C, umol.cm2.hr_uncorr, temp_c_level, temp_c_value) #keep only what we need

RespoR_PR <- RespoR_Normalized %>%
  dplyr::select(-umol.cm2.hr_uncorr, -Temp.C) %>% # remove to pivot
  pivot_wider(names_from = light_dark, values_from = umol.cm2.hr) %>% 
  rename(Respiration = DARK , NetPhoto = LIGHT) %>% # rename the columns
  mutate(Respiration = -1 * Respiration) %>%  # Make respiration positive
  mutate(GrossPhoto = Respiration + NetPhoto) %>% 
  pivot_longer(cols = Respiration:GrossPhoto, names_to = "PR", values_to = "Values") #values still in umol.cm2.hr


write_csv(RespoR_PR,here("Data","RespoFiles","TPC","PnR_rates.csv")) # export all the uptake rates
RespoR_PR <- read_csv(here("Data","RespoFiles","TPC","PnR_rates.csv"))
sp_keep <- c("Fcom","Prus", "Peyd", "Elam", "Maeq", "Tfro", "Ahya")
RespoR_PR <- RespoR_PR %>% filter(species %in% sp_keep)

#dev.off() # may need if plot doesn't run?
PR_plot <- RespoR_PR %>% 
  ggplot(aes(x = temp_c_value, y = Values, group=frag_ID, color = species)) +
  geom_point() +
  geom_line() +
  facet_wrap(~PR*species, scales = "free") +
  theme_bw() +
  labs(x = "Temperature °C", y = "umol.cm2.hr")+
  theme(strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"))

ggsave(here("Output", "TPC", "Graphs","PR_boxplots.pdf"),
       device = "pdf", height = 8, width = 8, PR_plot)

GP_plot <- RespoR_PR %>% filter(PR == "GrossPhoto") %>%
  ggplot(aes(x = temp_c_value, y = Values, group=frag_ID, color = species, shape = run_block)) +
  geom_point() +
  geom_line() +
  facet_wrap(~species, scales = "free") +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"))

NP_plot <- RespoR_PR %>% filter(PR == "NetPhoto") %>%
  ggplot(aes(x = temp_c_value, y = Values, group=frag_ID, color = species, shape = run_block)) +
  geom_point() +
  geom_line() +
  facet_wrap(~species, scales = "free") +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"))

Resp_plot <- RespoR_PR %>% filter(PR == "Respiration") %>%
  ggplot(aes(x = temp_c_value, y = Values, group=frag_ID, color = species, shape = run_block)) +
  geom_point() +
  geom_line() +
  facet_wrap(~species, scales = "free") +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"))

#######################
### making a df for just blank data for future use in plots ### 
Blank_only <- RespoR2 %>% 
  filter(blank == 1) %>% # grab the blanks
  group_by(temp_c_level, temp_c_value, run_block, light_dark) %>%
  #dplyr::select(blank.rate = umol.sec) %>% ## rename the blank column 
  summarise(blank.rate = mean(umol.sec, na.rm = TRUE))

#############################
write_csv(RespoR_Normalized , here("Data","RespoFiles","TPC","Respo_RNormalized_AllTPCRates.csv"))  
RespoR_Normalized <- read_csv(here("Data","RespoFiles","TPC","Respo_RNormalized_AllTPCRates.csv"))

## Plot the blanks across treatments to make sure nothing is funky
Blank_only %>%
  ggplot(aes(x = temp_c_value, blank.rate)) +
  geom_point()

#  basic plot of rates versus temp before you make the real TPC curve

RespoR_Normalized %>%
  ggplot(aes(x = temp_c_value, y = umol.cm2.hr, color = species))+
  geom_point() +
  facet_wrap(~light_dark)

##########################################################
### run an nls model for TPC curves for each species ###

# sharpeschoolhigh_1981
# https://padpadpadpad.github.io/rTPC/articles/rTPC.html

# first run Topt by species, then if not working, do it by FragID
# is I get an error, starting values are off, so play around with starting values that would make sense

# load packages
library(rTPC)
library(nls.multstart)
library(broom)
library(tidyverse)
library(here)

# read in data
df <- read_csv(here("Data","RespoFiles","TPC","PnR_rates.csv"))
df <- df %>% 
  mutate(id_block = paste(sample_ID,run_block,PR)) %>%
  mutate(full_ID = paste(sample_ID,run_block,temp_c_value, PR))
#log_df <- df %>% mutate(logValues = log(Values))

#remove data that is not biologically possible
#and remove run 4
bad_ids <- c("A01_TPC RUN1 24.5 NetPhoto","A01_TPC RUN1 26 NetPhoto", "A01_TPC RUN1 34 NetPhoto",
             "F08_TPC RUN6 34 NetPhoto",
             "A08_TPC RUN6 24.5 Respiration", "A08_TPC RUN6 24.5 NetPhoto", "A08_TPC RUN6 24.5 GrossPhoto",
             "C08_TPC RUN6 24.5 Respiration", "C08_TPC RUN6 24.5 NetPhoto", "C08_TPC RUN6 24.5 GrossPhoto")
bad_block <- c("B08_TPC RUN6 NetPhoto")
bad_run <- "RUN4"

df_no4 <- df %>%
  mutate(outlier_run = run_block %in% bad_run) %>% #remove run 4 for all 
  mutate(outlier_block = id_block %in% bad_block) %>% #remove a couple frags Run6
  mutate(outlier_ids = full_ID %in% bad_ids) %>% #remove a couple individuals for NP
  mutate(outlier_any = outlier_run | outlier_block | outlier_ids) 

#check graph of values to remove
ggplot(df_no4, aes(temp_c_value, Values, color = outlier_any)) +
  geom_point() +
  facet_wrap(PR ~ species, scales = "free_y") +
  theme_bw()

#drop outliers
df_clean <- df_no4 %>% filter(!outlier_any)
write_csv(df_clean, here("Data","RespoFiles","TPC","PnR_clean_no4.csv"))
df_clean <- read_csv(here("Data","RespoFiles","TPC","PnR_clean_no4.csv"))
#removes 289 data points (makes sense, one whole run plus quite a few)

#choose model for rTPC: 'sharpschoolhigh_1981'

#generate empty dataframes to fill in data for topt
topt_df <- tibble(rmax = as.numeric(),
                  topt = as.numeric(),
                  ctmin = as.numeric(),
                  ctmax = as.numeric(),
                  e = as.numeric(),
                  eh = as.numeric(),
                  q10 = as.numeric(),
                  thermal_safety_margin = as.numeric(),
                  thermal_tolerance = as.numeric(),
                  breadth = as.numeric(),
                  skewness = as.numeric(),
                  frag_ID = as.character(),
                  PR = as.character())

#and predicted data
preds_all <- tibble(
  PR = as.character(),
  frag_ID = as.character(),
  temp_c_value = as.numeric(),
  .fitted = numeric()
)

new_data <- tibble(temp_c_value = c(24.5, 26, 27, 28, 29, 30, 31, 32, 34))

for (j in unique(df_clean$PR)){
  pr = j
  PR_df <- df_clean %>%
    filter(PR == pr)
  
  for(i in unique(PR_df$frag_ID)){
    id = i
    my_df <- PR_df %>%
      filter(frag_ID == id)
    
    # get start vals
    start_vals <- get_start_vals(my_df$temp_c_value, my_df$Values, model_name = 'sharpeschoolhigh_1981')
    
    #issues with setting start values with some parameters
    #skip these
    if (anyNA(start_vals) || any(!is.finite(start_vals))) {
      message(sprintf("Skipping %s %s: could not set start values", id, pr))
      next
    }
    
    # get limits
    low_lims <- get_lower_lims(my_df$temp_c_value, my_df$Values, model_name = 'sharpeschoolhigh_1981')
    upper_lims <- get_upper_lims(my_df$temp_c_value, my_df$Values, model_name = 'sharpeschoolhigh_1981')
    
    # fit model
    fit <- nls_multstart(Values~sharpeschoolhigh_1981(temp = temp_c_value, r_tref,e,eh,th, tref = 27),
                         data = my_df,
                         iter = 500,
                         start_lower = start_vals - 10,
                         start_upper = start_vals + 10,
                         lower = low_lims,
                         upper = upper_lims,
                         supp_errors = 'Y')
    
    fit
    
    # calculate additional traits
    topt_params <- calc_params(fit) %>%
      # round for easy viewing
      mutate_all(round, 2) %>% 
      mutate(PR = pr,
             frag_ID = id)
    
    topt_df <- topt_df %>%
      rbind(topt_params)
    
    #generate predictions
    preds_id <- augment(fit, newdata = new_data) %>%
      transmute(PR = pr,
                frag_ID = id,
                temp_c_value = new_data$temp_c_value,
                .fitted = .fitted)
    
    preds_all <- bind_rows(preds_all, preds_id)
  }
}

#Code skips over a lot of the respiration values because it can't set start values
#these will print when you run
#skips no NP or GP

#add species names to predictions and topt dfs and pnr data
BioData <- read_csv(here("Data","RespoFiles","TPC","Fragment_Measurements_TPC.csv"))

preds_all <- preds_all %>% 
  dplyr::select(PR, frag_ID, temp_c_value, .fitted) %>% 
  left_join(BioData, by = "frag_ID")
topt_df <- topt_df %>% 
  dplyr::select(rmax, topt, ctmin, ctmax, e, eh, q10, thermal_safety_margin, thermal_tolerance, breadth, skewness, PR, frag_ID) %>% 
  left_join(BioData, by = "frag_ID")

#save data files
#clean = outliers removed
write_csv(topt_df, here("Data","RespoFiles","TPC","Topt_data_clean_no4.csv"))
write_csv(preds_all, here("Data","RespoFiles","TPC","Preds_data_clean_no4.csv"))

####Topt and other parameter graphs####
#read in dataframes and generate prediction dfs for each metric to graph
PnR_clean <- read_csv(here("Data","RespoFiles","TPC","PnR_clean_no4.csv"))
BioSp <- BioData %>% dplyr::select(frag_ID, full_species)
PnR_clean <-  PnR_clean %>% left_join(BioSp, by = "frag_ID")
sp_keep <- c("Fcom","Prus", "Peyd", "Elam", "Maeq", "Tfro", "Ahya")
PnR_clean <- PnR_clean %>% filter(species %in% sp_keep)
write_csv(PnR_clean, here("Data","RespoFiles","TPC","PnR_clean_no4_7sp.csv")) #P and R rate data cleaned
preds_all <- read_csv(here("Data","RespoFiles","TPC","Preds_data_clean_no4.csv"))
preds_all <- preds_all %>% filter(species %in% sp_keep)
write_csv(preds_all, here("Data","RespoFiles","TPC","Preds_data_clean_no4_7sp.csv")) #P and R rate data cleaned 7sp only
PnR_clean <- read_csv(here("Data","RespoFiles","TPC","PnR_clean_no4_7sp.csv"))
preds_all <- read_csv(here("Data","RespoFiles","TPC","Preds_data_clean_no4_7sp.csv"))
preds_gp <- preds_all %>% filter(PR == "GrossPhoto")
preds_np <- preds_all %>% filter(PR == "NetPhoto")
preds_resp <- preds_all %>% filter(PR == "Respiration")
#and topt data
topt_df <- read_csv(here("Data","RespoFiles","TPC","Topt_data_clean_no4.csv"))
topt_df <- topt_df %>% filter(species %in% sp_keep)
write_csv(topt_df, here("Data","RespoFiles","TPC","Topt_data_clean_no4_7sp.csv"))
topt_df <- read_csv(here("Data","RespoFiles","TPC","Topt_data_clean_no4_7sp.csv"))
topt_gp <- topt_df %>% filter(PR == "GrossPhoto")
topt_np <- topt_df %>% filter(PR == "NetPhoto")
topt_resp <- topt_df %>% filter(PR == "Respiration")

######plots of predicted TPCs with data#####

#gross photo
gp_pred_plot <- PnR_clean %>% filter(PR == "GrossPhoto") %>% ggplot(aes(temp_c_value, Values)) +
  geom_point(alpha = 0.7) +
  geom_line(data = preds_gp,
            aes(temp_c_value, .fitted, group = frag_ID),
            linewidth = 0.6, color = "darkblue") +
  theme_bw(base_size = 12) +
  theme(strip.text = element_text(face = "italic"))+
  #geom_vline(data = topt_gp,aes(xintercept = topt),linewidth = 0.3, color = "red") +
  #geom_hline(data = topt_gp,aes(yintercept = rmax),linewidth = 0.3, color = "darkgreen") +
  facet_wrap(~ full_species, scales = "free_y") +
  ylim(0.48,2.5) +
  labs(x = "Temperature (ºC)",
       y = expression("Gross Photosynthesis" ~ (mu*mol ~ cm^{-2} ~ h^{-1})),
       title = "Thermal performance: gross photosynthesis by species")

gp_pred_plot

ggsave(here("Output", "TPC", "Graphs", "gp_predicted_plot_7sp.pdf"),
       device = "pdf", height = 8, width = 8, gp_pred_plot)

#net photo
np_pred_plot <- PnR_clean %>% filter(PR == "NetPhoto") %>% ggplot(aes(temp_c_value, Values)) +
  geom_point(alpha = 0.7) +
  geom_line(data = preds_np,
            aes(temp_c_value, .fitted, group = frag_ID),
            linewidth = 0.6, color = "darkblue") +
  facet_wrap(~ full_species, scales = "free_y") +
  #geom_vline(data = topt_np,aes(xintercept = topt),linewidth = 0.3, color = "red") +
  #geom_hline(data = topt_np,aes(yintercept = rmax),linewidth = 0.3, color = "darkgreen") +
  theme_bw(base_size = 12) +
  theme(strip.text = element_text(face = "italic")) +
  ylim(0,1.6) +
  labs(x = "Temperature (ºC)",
       y = expression("Net Photosynthesis" ~ (mu*mol ~ cm^{-2} ~ h^{-1})),
       title = "Thermal performance: net photosynthesis by species")

np_pred_plot

ggsave(here("Output", "TPC", "Graphs","np_predicted_plot_7sp.pdf"),
       device = "pdf", height = 8, width = 8, np_pred_plot)

#respiration 
#this has been left alone for now since MP just focusing on GP and NP for poster
resp_pred_plot <- PnR_clean %>% filter(PR == "Respiration") %>% ggplot(aes(temp_c_value, Values)) +
  geom_point(alpha = 0.7) +
  geom_line(data = preds_resp, aes(temp_c_value, .fitted, group = frag_ID),
            linewidth = 0.6, color = "blue") +
  facet_wrap(~ species, scales = "free_y") +
  #geom_vline(data = topt_resp,aes(xintercept = topt),linewidth = 0.3, color = "red") +
  #geom_hline(data = topt_resp,aes(yintercept = rmax),linewidth = 0.3, color = "darkgreen") +
  theme_bw(base_size = 12) +
  labs(x = "Temperature (ºC)",
       y = "Respiration (umol.cm2.hr)",
       title = "Thermal performance: respiration by species")

resp_pred_plot


ggsave(here("Output", "TPC", "Graphs","resp_predicted_plot_clean_no4.pdf"),
       device = "pdf", height = 8, width = 6, resp_pred_plot)


#all rates stacked plots
PnR_gp <- PnR_clean %>% filter(PR == "GrossPhoto")
species_gp_plot <- PnR_gp %>% ggplot(aes(temp_c_value, Values, color = full_species)) +
  geom_point(alpha = 0.7) +
  geom_line(data = preds_gp,
            aes(temp_c_value, .fitted, group = frag_ID, color = full_species),
            linewidth = 0.6) +
  #facet_wrap(~ PR, scales = "free_y") +
  scale_color_discrete(labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  theme_bw(base_size = 12) +
  labs(x = "Temperature (ºC)",
       y = expression("Gross Photosynthesis" ~ (mu*mol ~ cm^{-2} ~ h^{-1})),
       title = "Thermal performance",
       color = "Species")

species_gp_plot

ggsave(here("Output", "TPC", "Graphs", "gp_stacked_plot_7sp.pdf"),
       device = "pdf", height = 8, width = 8, species_gp_plot)

PnR_np <- PnR_clean %>% filter(PR == "NetPhoto")
species_np_plot <- PnR_np %>% ggplot(aes(temp_c_value, Values, color = full_species)) +
  geom_point(alpha = 0.7) +
  geom_line(data = preds_np,
            aes(temp_c_value, .fitted, group = frag_ID, color = full_species),
            linewidth = 0.6) +
  #facet_wrap(~ PR, scales = "free_y") +
  scale_color_discrete(labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  theme_bw(base_size = 12) +
  labs(x = "Temperature (ºC)",
       y = expression("Net Photosynthesis" ~ (mu*mol ~ cm^{-2} ~ h^{-1})),
       title = "Thermal performance",
       color = "Species")

species_np_plot

ggsave(here("Output", "TPC", "Graphs", "np_stacked_plot_7sp.pdf"),
       device = "pdf", height = 8, width = 8, species_np_plot)

##########################################################
####stats to look at differences in thermal performance metrics####
#interested in the effect of species (species) on these parameters:
#rmax, topt, e, breadth
#include random effect of genotype (frag_ID)

#load libraries
library(here)
library(tidyverse)
library(lme4)
library(lmerTest)
library(emmeans)
library(performance)
library(DHARMa)

sp_cols <- c(
  "Acropora hyacinthus" = '#d8aedd',
  "Echinopora lamellosa" = '#ba7999',
  "Favites complanata"   = '#dd4124',
  "Montipora aequituberculata" = '#ed8b00',
  "Montipora vietnamensis" = '#efbc82',
  "Pachyseris rugosa" = '#edd746',
  "Pocillopora eydouxi" = '#d0e2af',
  "Porites cylindrica" = '#45681e',
  "Porites rus" = '#7bbcd5',
  "Turbinaria frondens" = '#00496f'
)

#read in data
topt_df <- read_csv(here("Data","RespoFiles","TPC","Topt_data_clean_no4.csv"))
topt_df <- topt_df %>% filter(sample_ID != "B08_TPC")

#convert to long so you can plot everything at once
metrics <- c("topt", "rmax", "e", "breadth") #select metrics
grouping_vars <- c("species", "PR", "frag_ID", "full_species") #select grouping variables

topt_long <- topt_df %>% #make long dataframe
  dplyr::select(all_of(c(grouping_vars, metrics))) %>%
  pivot_longer(cols = all_of(metrics),
               names_to = "metric",
               values_to = "value")

#now generate mean and se dataframe
se_fun <- function(x) {
  n <- sum(!is.na(x))
  if (n <= 1) return(NA_real_)
  sd(x, na.rm = TRUE) / sqrt(n)
}

topt_summary <- topt_long %>%
  group_by(full_species,PR, metric) %>%
  summarise(
    n    = sum(!is.na(value)),
    mean = mean(value, na.rm = TRUE),
    se   = se_fun(value),
    .groups = "drop")

#first visualize data - plots of Topt and other variables from TPCs######
topt_plot <- ggplot() +
  geom_jitter(data = topt_long,
              aes(x = full_species, y = value, color = full_species),
              width = 0.15, alpha = 0.8) +
  geom_errorbar(data = topt_summary,
                aes(x = full_species,
                    ymin = mean - se, ymax = mean + se),
                width = 0.2, linewidth = 0.6) +
  geom_point(data = topt_summary,
             aes(x = full_species, y = mean),
             size = 2) +
  facet_wrap(PR ~ metric, scales = "free_y") +
  theme_bw(base_size = 12) +
  scale_color_manual(values = sp_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))
  #labs(x = )
topt_plot

ggsave(here("Output", "TPC", "Graphs","Topt_allparams_clean_no4.pdf"),
       device = "pdf", height = 8, width = 8, topt_plot)

#individual dataframes to put in individual plots so they can be ordered (live laugh love i guess)
rmax_mean_gp <- topt_summary %>% filter(PR == "GrossPhoto") %>% filter(metric == "rmax")
topt_rmax_gp <- topt_long %>% 
  filter(PR == "GrossPhoto") %>% 
  filter(metric == "rmax") %>%
  left_join(rmax_mean_gp, by = "full_species") %>% 
  mutate(full_species = fct_reorder(full_species, mean))

rmax_mean_np <- topt_summary %>% filter(PR == "NetPhoto") %>% filter(metric == "rmax")
topt_rmax_np <- topt_long %>% 
  filter(PR == "NetPhoto") %>% 
  filter(metric == "rmax") %>%
  left_join(rmax_mean_np, by = "full_species") %>% 
  mutate(full_species = fct_reorder(full_species, mean))

topt_mean_gp <- topt_summary %>% filter(PR == "GrossPhoto") %>% filter(metric == "topt")
topt_topt_gp <- topt_long %>% 
  filter(PR == "GrossPhoto") %>% 
  filter(metric == "topt") %>%
  left_join(topt_mean_gp, by = "full_species") %>% 
  mutate(full_species = fct_reorder(full_species, mean))

topt_mean_np <- topt_summary %>% filter(PR == "NetPhoto") %>% filter(metric == "topt")
topt_topt_np <- topt_long %>% 
  filter(PR == "NetPhoto") %>% 
  filter(metric == "topt") %>%
  left_join(topt_mean_np, by = "full_species") %>% 
  mutate(full_species = fct_reorder(full_species, mean))

e_mean_gp <- topt_summary %>% filter(PR == "GrossPhoto") %>% filter(metric == "e")
topt_e_gp <- topt_long %>% 
  filter(PR == "GrossPhoto") %>% 
  filter(metric == "e") %>%
  left_join(e_mean_gp, by = "full_species") %>% 
  mutate(full_species = fct_reorder(full_species, mean))

e_mean_np <- topt_summary %>% filter(PR == "NetPhoto") %>% filter(metric == "e")
topt_e_np <- topt_long %>% 
  filter(PR == "NetPhoto") %>% 
  filter(metric == "e") %>%
  left_join(e_mean_np, by = "full_species") %>% 
  mutate(full_species = fct_reorder(full_species, mean))

breadth_mean_gp <- topt_summary %>% filter(PR == "GrossPhoto") %>% filter(metric == "breadth")
topt_breadth_gp <- topt_long %>% 
  filter(PR == "GrossPhoto") %>% 
  filter(metric == "breadth") %>%
  left_join(breadth_mean_gp, by = "full_species") %>% 
  mutate(full_species = fct_reorder(full_species, mean))

breadth_mean_np <- topt_summary %>% filter(PR == "NetPhoto") %>% filter(metric == "breadth")
topt_breadth_np <- topt_long %>% 
  filter(PR == "NetPhoto") %>% 
  filter(metric == "breadth") %>%
  left_join(breadth_mean_np, by = "full_species") %>% 
  mutate(full_species = fct_reorder(full_species, mean))

topt_breadth_plot <- ggplot() +
  geom_jitter(data = topt_breadth_np, aes(x = full_species, y = value, color = full_species), width = 0.15, alpha = 0.8) +
  geom_errorbar(data = breadth_mean_np,aes(x = full_species, ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 0.6) +
  geom_point(data = breadth_mean_np, aes(x = full_species, y = mean), size = 2) +
  theme_bw(base_size = 22) +
  stat_summary(data = topt_breadth_np, aes(x = full_species, y = value), geom = "text", fun = max, vjust = -0.5, size = 8,
               label = c("a", "ab", "ab", "ab", "ab", "ab", "ab", "ab", "ab", "b"))+
  theme(axis.text.x = element_blank(), axis.title.x = element_blank(),legend.position = "right")+
  #theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"), legend.position = "none")+
  scale_color_manual(values = sp_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  labs(x = "Species", y = "Breadth (°C)", color = "Species")

topt_e_plot <- ggplot() +
  geom_jitter(data = topt_e_np, aes(x = full_species, y = value, color = full_species), width = 0.15, alpha = 0.8) +
  geom_errorbar(data = e_mean_np,aes(x = full_species, ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 0.6) +
  geom_point(data = e_mean_np, aes(x = full_species, y = mean), size = 2) +
  theme_bw(base_size = 22) +
  ylim(0,1)+
  stat_summary(data = topt_e_np, aes(x = full_species, y = value), geom = "text", fun = max, vjust = -0.5, size = 8,
               label = c("a", "ab", "ab", "ab", "ab", "ab", "ab", "ab", "ab", "b"))+
  theme(axis.text.x = element_blank(), axis.title.x = element_blank(),legend.position = "right")+
  #theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"), , legend.position = "none", axis.title.y = element_text(face = "italic"))+
  scale_color_manual(values = sp_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  labs(x = "Species", y = "E (eV)", color = "Species")

topt_topt_plot <- ggplot() +
  geom_jitter(data = topt_topt_np, aes(x = full_species, y = value, color = full_species), width = 0.15, alpha = 0.8) +
  geom_errorbar(data = topt_mean_np,aes(x = full_species, ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 0.6) +
  geom_point(data = topt_mean_np, aes(x = full_species, y = mean), size = 2) +
  theme_bw(base_size = 22) +
  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), legend.position = "right")+
  #theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"), legend.position = "none")+
  scale_color_manual(values = sp_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  labs(x = "Species", y = "Thermal optimum (°C)", color = "Species")

topt_rmax_plot <- ggplot() +
  geom_jitter(data = topt_rmax_np, aes(x = full_species, y = value, color = full_species), width = 0.15, alpha = 0.8) +
  geom_errorbar(data = rmax_mean_np,aes(x = full_species, ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 0.6) +
  geom_point(data = rmax_mean_np, aes(x = full_species, y = mean), size = 2) +
  theme_bw(base_size = 22) +
  stat_summary(data = topt_rmax_np, aes(x = full_species, y = value), geom = "text", fun = max, vjust = -0.5, size = 8,
               label = c("a", "ab", "ab", "ab", "ab", "ab", "ab", "ab", "b", "b"))+
  theme(axis.text.x = element_blank(), axis.title.x = element_blank(),legend.position = "right")+
  #theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "italic"), legend.position = "none")+
  scale_color_manual(values = sp_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  ylim(0,2)+
  labs(x = "Species", y = expression("Rmax" ~ (mu*mol ~ cm^{-2} ~ h^{-1})), , color = "Species")

library(ggpubr)
np_topt_plots <- ggarrange(topt_rmax_plot, topt_topt_plot, 
                           nrow = 1, ncol = 2, legend = "right", common.legend = TRUE)
np_topt_plots

ggsave(here("Output","TPC","Graphs","np_topt_plots_2.pdf"), np_topt_plots, h = 5, w = 15, dpi = 300)


#topt plot all together
topt_plot <- ggplot() +
  geom_jitter(data = topt_long,
              aes(x = full_species, y = value, color = full_species),
              width = 0.15, alpha = 0.8) +
  geom_errorbar(data = topt_summary,
                aes(x = full_species,
                    ymin = mean - se, ymax = mean + se),
                width = 0.2, linewidth = 0.6) +
  geom_point(data = topt_summary,
             aes(x = full_species, y = mean),
             size = 2) +
  facet_wrap(PR ~ metric, scales = "free_y") +
  theme_bw(base_size = 12) +
  scale_color_manual(values = sp_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))
#labs(x = )
topt_plot



#check distributions of data as well
ggplot(topt_long, aes(value)) +
  geom_histogram(bins = 30) + 
  facet_wrap(PR~metric, scales = "free")

#run a loop to model the data and get the output and the pairwise comparisons
library(broom)

PR_levels <- c("GrossPhoto", "NetPhoto", "Respiration")
resp_metrics <- c("rmax", "topt", "e", "breadth")

#data lists
emm_list   <- list()
emm_pair_list <- list()
anova_list <- list()

for (pr in PR_levels) {
  dat_pr <- topt_long %>% filter(PR == pr)
  
  for (resp in resp_metrics) {
    dat <- dat_pr %>% filter(metric == resp)
    
    fit <- lm(value ~ species, data = dat)
    
    # ANOVA table -> tidy tibble
    an_tbl <- broom::tidy(anova(fit)) %>%
      mutate(PR = pr, metric = resp, .before = 1)
    anova_list[[paste(pr, resp, sep = "__")]] <- an_tbl
    
    # EMMs per species (+ CIs) -> tibble
    emm_obj <- emmeans::emmeans(fit, ~ species)
    emm_pairs <- pairs(emm_obj)
    emm_pair_tbl <- as_tibble(summary(emm_pairs, infer = TRUE)) %>%
      mutate(PR = pr, metric = resp, .before = 1)
    emm_tbl <- as_tibble(summary(emm_obj, infer = TRUE)) %>%
      mutate(PR = pr, metric = resp, .before = 1)
    emm_list[[paste(pr, resp, sep = "__")]] <- emm_tbl
    emm_pair_list[[paste(pr, resp, sep = "__")]] <- emm_pair_tbl
  }
}

#add values to table
emm_all   <- bind_rows(emm_list)
anova_all <- bind_rows(anova_list)
emm_pair_set <- bind_rows(emm_pair_list)

#NP


write_csv(emm_all, here("Data","RespoFiles","TPC","emmeans_all_PR_metrics.csv"))
write_csv(anova_all, here("Data","RespoFiles","TPC","anova_all_PR_metrics.csv"))

emm_plot <- ggplot(emm_all, aes(x = emmean, y = species, color = species)) +
  geom_point() +
  geom_errorbar(aes(xmin = lower.CL, xmax = upper.CL), width = 0.2) +
  theme_bw(base_size = 12) +
  facet_wrap(PR ~ metric, scales = "free") +
  labs(title = "emmeans + confidence intervals")
emm_plot

ggsave(here("Output", "TPC", "Graphs","Emmeans_allparams_clean_no4.pdf"),
       device = "pdf", height = 8, width = 8, emm_plot)

### Compare with Danielle's Data ###
PnR_clean <- read_csv(here("Data","RespoFiles","TPC","PnR_clean_no4.csv"))

species_cols <- c(
  "Ahya" = '#d8aedd',
  "Elam" = '#ba7999',
  "Fcom"   = '#dd4124',
  "Maeq" = '#ed8b00',
  "Mvie" = '#efbc82',
  "Prug" = '#edd746',
  "Peyd" = '#d0e2af',
  "Pcyl" = '#45681e',
  "Prus" = '#7bbcd5',
  "Tfro" = '#00496f'
)

se_fun <- function(x) {
  n <- sum(!is.na(x))
  if (n <= 1) return(NA_real_)
  sd(x, na.rm = TRUE) / sqrt(n)
}

PnR_28 <- PnR_clean %>% filter(temp_c_value == 28)
PnR_28_summary <- PnR_28 %>%
  group_by(species, PR) %>%
  summarise(n = sum(!is.na(Values)),
            mean = mean(Values, na.rm = TRUE),
            se = se_fun(Values),
            .groups = "drop")

GP_28 <- PnR_28 %>% filter(PR == "GrossPhoto")
NP_28 <- PnR_28 %>% filter(PR == "NetPhoto")
R_28 <- PnR_28 %>% filter(PR == "Respiration")

GP_31 <- PnR_31 %>% filter(PR == "GrossPhoto")
NP_31 <- PnR_31 %>% filter(PR == "NetPhoto")
R_31 <- PnR_31 %>% filter(PR == "Respiration")

sp_mod <- lm(Values~species, data = GP_31)
Anova(sp_mod)
#summary(sp_mod)
#check_model(sp_mod)

emm_obj <- emmeans::emmeans(sp_mod, ~ species)
emm_pairs <- pairs(emm_obj)
emm_pairs

#GP 28 - Maeq lower than Elam, Fcom, Pcyl, and Prus
#Elam - Maeq  0.60595 0.174 40   3.482  0.0358
#Fcom - Maeq  0.65241 0.174 40   3.749  0.0178
#Maeq - Pcyl -0.59749 0.174 40  -3.433  0.0405
#Maeq - Prus -0.67170 0.174 40  -3.860  0.0132

#R 28 - Maeq lower than Prus and Pcyl
# Maeq - Prus -0.23000 0.0637 40  -3.612  0.0256
# Maeq - Pcyl -0.24705 0.0637 40  -3.880  0.0125

#GP 31 - 
# Fcom - Maeq  7.64e-01 0.184 40   4.150  0.0058
# Maeq - Pcyl -6.18e-01 0.184 40  -3.358  0.0488
# Maeq - Prug -6.18e-01 0.184 40  -3.358  0.0488
#Maeq - Prus -7.32e-01 0.184 40  -3.979  0.0095


#R 31 - Maeq only lower than Prus (almost Pcyl)
# Maeq - Prus -0.28185 0.0718 40  -3.927  0.0109
# Maeq - Pcyl -0.23840 0.0718 40  -3.322  0.0533

PnR_28_ordered <- PnR_28 %>% 
  left_join(PnR_28_summary) %>% 
  group_by(PR) %>%
  mutate(species = fct_reorder(species, mean))  # ascending by mean

PnR_28_plot <- ggplot() +
  geom_jitter(data = PnR_28_ordered, aes(x = species, y = Values, color = species), width = 0.15, alpha = 0.8) +
  facet_wrap(.~PR)+
  #stat_summary(data = PnR_28_ordered, aes(x = species, y = Values), geom = "text", fun = max, vjust = -0.5, size = 8,
  #             label = c("a", "a", "a", "a", "ab", "ab", "b"))+
  geom_errorbar(data = PnR_28_summary, aes(x = species, ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 0.6) +
  geom_point(data = PnR_28_summary, aes(x = species, y = mean), size = 2) +
  theme_bw(base_size = 22) +
  theme(legend.position = "right", axis.text.x = element_blank(), axis.title.x = element_blank()) +
  #theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust=1, face = "italic")) +
  scale_color_manual(values = species_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  #ylim(25,325)+
  labs(x = "Species", color = "Species", y = expression("28 °C Metabolic Rate" ~ (mu*mol ~ cm^{-2} ~ h^{-1})))
PnR_28_plot

ggsave(here("Output", "TPC", "Graphs","PnR_means_28C.pdf"), device = "pdf", height = 8, width = 12, PnR_28_plot)

PnR_31 <- PnR_clean %>% filter(temp_c_value == 31)
PnR_31_summary <- PnR_31 %>%
  group_by(species, PR) %>%
  summarise(n = sum(!is.na(Values)),
            mean = mean(Values, na.rm = TRUE),
            se = se_fun(Values),
            .groups = "drop")

PnR_31_ordered <- PnR_31 %>% 
  left_join(PnR_31_summary) %>% 
  group_by(PR)%>%
  mutate(species = fct_reorder(species, mean))  # ascending by mean

PnR_31_plot <- ggplot() +
  geom_jitter(data = PnR_31_ordered, aes(x = species, y = Values, color = species), width = 0.15, alpha = 0.8) +
  facet_wrap(.~PR)+
  #stat_summary(data = PnR_31_ordered, aes(x = species, y = Values), geom = "text", fun = max, vjust = -0.5, size = 8,
  #             label = c("a", "a", "a", "a", "ab", "ab", "b"))+
  geom_errorbar(data = PnR_31_summary, aes(x = species, ymin = mean - se, ymax = mean + se), width = 0.2, linewidth = 0.6) +
  geom_point(data = PnR_31_summary, aes(x = species, y = mean), size = 2) +
  theme_bw(base_size = 22) +
  theme(legend.position = "right", axis.text.x = element_blank(), axis.title.x = element_blank()) +
  #theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust=1, face = "italic")) +
  scale_color_manual(values = species_cols, labels = function(x) parse(text = paste0("italic('", gsub("'", "\\\\'", x), "')")))+
  #ylim(25,325)+
  labs(x = "Species", color = "Species", y = expression("31 °C Metabolic Rate" ~ (mu*mol ~ cm^{-2} ~ h^{-1})))
PnR_31_plot

ggsave(here("Output", "TPC", "Graphs","PnR_means_31C.pdf"), device = "pdf", height = 8, width = 12, PnR_31_plot)


